Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2014-08-13 writeups/dissertation/dissertation.tex
foo

Another issue with LSA is that adding additional observations to an previously-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
Since computing the SVD is the most computationally expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.

@@

Random permutations uses one of simplest permutations possible to create a new environment vector.
This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
Further, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.

@@

Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{http://tex.stackexchange.org} and \url{http://superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.

@@

It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
This suggests that more cognitively-plausible memory retrieval models may outperform more general machine learning approaches in this particular environment
(i.e., when predicting the user-chosen tags for human-created content on social media sites).

@@

A tag suggestion tool has been deployed for the \url{http://tex.stackexchange.org} \parencite{LatexTags2013} and \url{http://superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
Latex, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
An example of the interface for a created post on the SuperUser site is included in Figure \ref{figSuperUserSuggestion}.

@@

A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
This algorithm was specifically created for tokenizing Twitter content.
However, during testing it was also found to be robust and well suited for StackOverflow content.
The initial model results with the new tokenizer were qualitatively very similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
This is most likely because the language for the samples taken for both sites is informal English,
and it seams reasonable to think that tokenizing all of the informal language and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar for both algorithms on StackOverflow,
and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.

@@

The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
What is interesting is the large tag space for these users.
There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
This helps illustrate how challenging modeling this task can be.
The model's goal is to predict at each instance what tag will be chosen, but since the tag space is so large and varied, the model must predict the correct few tags that are used from a much larger set of possible tags.

@@

The best-fit values are close to the \num{0.5} default value used in ACT-R when the optimized learning form is used.
Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
Comparing best-fit values will be more thoroughly tested when looking at the values across all dataset slices.

>>>>

Another issue with LSA is that adding additional observations to a previously-computed reduced representation requires rerunning the SVD \parencite{Farahat2004}.
Since computing the SVD is the most computationally expensive operation with LSA, it is difficult to see how LSA could be implemented in incremental domains where information is periodically added.
For Twitter and StackOverflow, it might certainly be useful to update the strength of association representation when new information arrives, especially for Twitter, where hashtag use and news content change rapidly.

@@

Random permutations uses one of simplest operations possible to create a new environment vector.
This operation is much less costly than circular convolution used by the BEAGLE model, which allows random permutations to scale better and handle larger datasets.
Further, even though random permutations is a simpler representation than BEAGLE, it also performs better when trained on the same sized corpus.

@@

Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.

@@

It is interesting that in both studies the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
This suggests that more cognitively-plausible memory retrieval models may outperform more general machine learning approaches in this particular environment
(i.e., when predicting the user-chosen tags for human-created content on social media sites).

@@

A tag suggestion tool has been deployed for the \url{tex.stackexchange.org} \parencite{LatexTags2013} and \url{superuser.com} \parencite{SuperUserTags2013} StackExchange sites.
Latex, SuperUser, and StackOverflow are all part of the broader StackExchange network of question and answer sites, and they provide the same user interface and workflow when asking questions.
For the two sites where a tag suggestion tool has been deployed, once a user has created a question, suggested tags appear just below the text field where the user types the tags to associate with the question.
The post author is not forced to use the suggested tags, and is free to choose tags from them or tags not represented in the suggested set.
An example of the interface for a created post on the SuperUser site is included in Figure \ref{figSuperUserSuggestion}.

@@

A custom tokenization algorithm created by \textcite{Owoputi2013} was used to chunk the words in the text.
This algorithm was specifically created for tokenizing Twitter content.
However, during testing it was also found to be robust and well suited for StackOverflow content.
The initial model results with the new tokenizer were qualitatively very similar to the model results found in \textcite{Stanley2013} where the Python Natural Language Processing (NLP) toolkit \parencite{Bird2009} was used.
This is most likely because the language for the samples taken for both sites is informal English,
and it seams reasonable to think that tokenizing all of the informal language and characters included in Twitter is a much more difficult task than tokenizing most of the words used on StackOverflow.
So the same tokenization algorithm was used for both StackOverflow and Twitter, since performance was similar to \textcite{Stanley2013} for StackOverflow,
and it seemed beneficial to use the same algorithm for both datasets to make comparisons across datasets more equitable.

@@

The specific value of the hashtag used for both plots is a nominal variable and not important for this analysis.
What is interesting is the large tag space for these users.
There are bands of commonly-used tags for each user, but there are also many instances where new tags are used, and a large percentage of tags used are not in the most common bands.
This helps illustrate how challenging modeling this task can be.
The model's goal is to predict at each instance what tag will be chosen, but to do so the model must predict the correct few tags from a much larger set of tags whose frequency of occurrence is dynamically changing over time.

@@

The best-fit values are close to the \num{0.5} default value set in ACT-R when the optimized learning form is used.
Comparing across optimized learning and the standard form, the best-fit values for the standard form are slightly higher for both the StackOverflow and Twitter dataset slices.
Comparing best-fit values will be more thoroughly tested when looking at the values across all dataset slices.
